The first high-redshift Changing Look Quasars

### 2019-11-07 tex/other_tex/201909/CIV_notes_20190920v2.tex
CIV tex/ tidy up

In this paper we use the wavelengths of 1548.202 and 1550.774 \AA\ for
the \civ doublet \citep{Kramida2018}.
%%  The singly ionized hydrogen atom H+ is referred to as HII, doubly ionized oxygen O++ is called OIII, etc.
%%  “24.4 eV is energy needed for C I to get rid of two electrons, and become doubly-ionised C III”
%%  https://en.wikipedia.org/wiki/Ionization_energies_of_the_elements_(data_page)
The ground state of carbon is 1$s^2$ 2$s^2$ 2$p^2$.  Using the
\href{https://physics.nist.gov/PhysRefData/ASD/ionEnergy.html}{NIST
Atomic Spectra Database Ionization Energies Form} for ionisation
energies, we see 11.3 eV is the energy needed for C I to dislodge one
electron and become singly-ionised C II; 24.4 eV is energy needed for
singly-ionised C II to dislodge an additional electron, and become
doubly-ionised C III, and 47.89 eV (3.519 Ry) is required for
doubly-ionised C III become triply--ionised \civ.  64.49 eV (4.74 Ry)
is the energy needed to ionize \civ itself.
%%
The 1548.202 and 1550.774 \AA\ emission doublet is created by the 2$p$
$^{2}P_{0}$ $---$ 2$s$ $^{2}S$ transition with energies 64,484
cm$^{-1}$ and 64591 cm$^{-1}$ \citep[e.g.][]{Moore1993}.
%Ground  Shells:  1$s^{2}$2$s$. 
%Ground  Level: $^{2}S_{1/2}$. 
%Ionized  Level: 1$s^{2}$ $^{1}S_{0}$.

%%\subsection{Literature Search}
\civ\ variability has been long studied, \citep[e.g., ][]{Baldwin1977,
Gaskell1982, Gregory1982, Wilkes1986, Espey1989, Espey1990Erratum,
ZhengSulentic1990, Corbin1990, Corbin1991, Weymann1991,
Dimitrijevic1992, TytlerFan1992, Wills1993, Brotherton1994, Osmer1994,
Laor1995, McIntosh1999}.
%%
%\citet{Gregory1982} reported time variability in both Lya and \civ 1550 emission lines in the Seyfert galaxy NGC 5548.
%
%\citet{1994}
%\includegraphics[]{../../../../../Users/npr1/Downloads/1994ASPC___69____1P.pdf}

%% NAZAROVA 2003
%% BROAD EMISSION LINES La, C IV AND Hb IN NGC 5548
%% Astronomical and Astrophysical Transactions
%% Vol. 22, Nos. 4–5, August–October 2003, pp. 681–689
\citep{Nazarova2003} ...

\citet{Wilhite2006} examine the variability of a C~\textsc{iv} 
sample of 105 quasars observed at multiple epochs by the SDSS. 
They find a strong correlation between the change in the C~\textsc{iv} line flux and the
change in the line width, but no correlations between the change in 
flux and changes in line center and skewness.  These authors find 
that the relation between line flux change and line width change is consistent with a model in
which a broad line base varies with greater amplitude than the line
core. The C~\textsc{iv} lines in these 
high-luminosity quasars appear to be less responsive to continuum
variations than those in lower luminosity AGN.  
%% Additionally, we find no evidence for variability of the well known
blueshift of the C~\textsc{iv} line with respect to the low-ionization
Mg~\textsc{ii}$\lambda$2798 line in the highest flux objects,
indicating that this blueshift might be useful as a measure of
orientation.

%% Punsly 2010
%% Astrophysical Journal, 713, 232
%% THE REDSHIFTED EXCESS IN QUASAR C iv BROAD EMISSION LINES

\citet{Richards2011} explored the BELR in over 30,000 $z > 1.54$ SDSS
quasars, concentrating on the properties of the \civ\ emission
line. We consider two well-known effects involving the \civ\ emission
line: the anti-correlation between the \civ\ EQW and luminosity (i.e.,
the BEff) and the blueshifting of the peak of \civ\ emission with
respect to the systemic redshift. These authors conclude that these
two \civ\ parameters (EQW and blueshift) are capturing an important
trade-off between ``disk'' and ``wind'' components in the disk-wind
model of accretion disks \citep[e.g.,][]{Murray1995, Elvis2000, Proga2000}.  with one dominating over the other depending on the
shape of the SED \citep[][]{Leighly2004b}.  (strong \civ EQW indicates
a more ionizing SED and large \civ blueshift indicating a less
ionizing SED).

%% \citet{Grier2015}....
%% \includegraphics[]{../../../../../Users/npr1/Downloads/Grier_2015_ApJ_806_111.pdf}
%% \citet{Denney2015}, 
%%Denney, Kelly; SDSS-RM Team, ``The Effects of S/N on Measuring CIV Broad Emission Line Widths in Quasars - An Early Science Result from the Sloan Digital Sky Survey Reverberation Mapping Project'' ; http://adsabs.harvard.edu/abs/2015AAS...22520403D
The Sloan Digital Sky Survey Reverberation Mapping (SDSS-RM) Project 
\citep{Shen2015} has a  monitored $\sim$350 quasars with \civ
Noting the biases associated with \civ Emission Line Properties 
\citep[e.g. increasing systematic offsets with decreasing signal-to-noise][]{Denney2016}, 
\citet{Grier2019} report 
report significant time delays between the continuum and the CIV 1549 emission line in 52 quasars, 
and investigate the \civ radius-luminosity relationship.

%% \citet{Coatman2017} 
%The \civ $\lambda \lambda$1498,1501 broad emission line is visible in optical spectra to redshifts exceeding $z\sim5$. 
\civ has long been known to exhibit significant displacements to the blue and these `blueshifts' almost certainly signal the presence of strong outflows. As a consequence, single-epoch virial black hole (BH) mass estimates derived from \civ velocity widths are known to be systematically biased compared to masses from the hydrogen Balmer lines. 
\citet{Coatman2017} use a large sample of 230 high-luminosity (LBol = 1045.5-1048 erg s-1), redshift $1.5 < z < 4.0$ quasars with both \civ and Balmer line spectra, we have quantified the bias in \civ BH masses as a function of the \civ blueshift. \civ BH masses are shown to be a factor of 5 larger than the corresponding Balmer-line masses at C IV blueshifts of 3000 km s$^{-1}1$ and are overestimated by almost an order of magnitude at the most extreme blueshifts, $\gtrsim$5000 km s$^{-1}$. 
%Using the monotonically increasing relationship between the \civ blueshift and the mass ratio BH(\civ)/BH(H$\alpha$), we derive an empirical correction to all C IV BH masses. The scatter between the corrected C IV masses and the Balmer masses is 0.24 dex at low C IV blueshifts ($\sim$0 km s-1) and just 0.10 dex at high blueshifts ($\sim$3000 km s$^{-1}$), compared to 0.40 dex before the correction. The correction depends only on the C IV line properties - i.e. full width at half-maximum and blueshift - and can therefore be applied to all quasars where C IV emission line properties have been measured, enabling the derivation of unbiased virial BH-mass estimates for the majority of high-luminosity, high-redshift, spectroscopically confirmed quasars in the literature.

@@

\subsection{Line Properties}. 
We use the measured quasar emission line properties from \citet{Shen2011} and \citet{Hamann2017}. 
\citet{Hamann2017} in particular investigate in robust detail the UV continuum and the \civ (and \nv $\lambda$ $\lambda$1238, 1242) emission lines in 
over 200,000 quasars in BOSS DR12Q \citep{Paris2017}\footnote{This emission-line catalog 
can be downloaded from \href{https://datadryad.org/stash/dataset/doi:10.6086/D1H59V}{here}.}
The quasar redshift are limited to the range 1.53 ≤ ze ≤ 5.0 so that C IV and the adjacent continuum are covered by BOSS. 
These measurements provide line profile information and N V/C IV flux ratios

>>>>

In this paper we use the wavelengths of 1548.202 and 1550.774 \AA\ for the \civ doublet \citep{Kramida2018}. The ground state of carbon is 1$s^2$ 2$s^2$ 2$p^2$.  Using the \href{https://physics.nist.gov/PhysRefData/ASD/ionEnergy.html}{NIST Atomic Spectra Database Ionization Energies Form} for ionisation energies, we see 11.3 eV is the energy needed for C I to dislodge one electron and become singly-ionised C II; 24.4 eV is energy needed for singly-ionised C II to dislodge an additional electron, and become doubly-ionised C III, and 47.89 eV (3.519 Ry) is required for doubly-ionised C III become triply--ionised \civ.  64.49 eV (4.74 Ry) is the energy needed to ionize \civ itself. The 1548.202 and 1550.774 \AA\ emission doublet is created by the 2$p$ $^{2}P_{0}$ $-$ 2$s$ $^{2}S$ transition with energies 64,484 cm$^{-1}$ and 64,591 cm$^{-1}$ \citep[e.g.][]{Moore1993}.

%%\subsection{Literature Search}
\civ\ variability has been long studied, \citep[e.g., ][]{Baldwin1977,
Gaskell1982, Gregory1982, Wilkes1986, Espey1989, Espey1990Erratum,
ZhengSulentic1990, Corbin1990, Corbin1991, Weymann1991,
Dimitrijevic1992, TytlerFan1992, Wills1993, Brotherton1994, Osmer1994,
Laor1995, McIntosh1999, Nazarova2003}.
%%
\citet{Wilhite2006} examine the variability of a C~\textsc{iv} sample
of 105 quasars observed at multiple epochs by the SDSS.  They find a
strong correlation between the change in the C~\textsc{iv} line flux
and the change in the line width, but no correlations between the
change in flux and changes in line center and skewness.  These authors
find that the relation between line flux change and line width change
is consistent with a model in which a broad line base varies with
greater amplitude than the line core. The C~\textsc{iv} lines in these
high-luminosity quasars appear to be less responsive to continuum
variations than those in lower luminosity AGN.  \citet{Wilhite2006}
find no evidence for variability of the well known blueshift of the
C~\textsc{iv} line with respect to the low-ionization
Mg~\textsc{ii}$\lambda$2798 line in the highest flux objects.

\citet{Richards2011} explored the BELR in over 30,000 $z > 1.54$ SDSS
quasars, concentrating on the properties of the \civ\ emission
line. We consider two well-known effects involving the \civ\ emission
line: the anti-correlation between the \civ\ EQW and luminosity (i.e.,
the BEff) and the blueshifting of the peak of \civ\ emission with
respect to the systemic redshift. These authors conclude that these
two \civ\ parameters (EQW and blueshift) are capturing an important
trade-off between ``disk'' and ``wind'' components in the disk-wind
model of accretion disks \citep[e.g.,][]{Murray1995, Elvis2000,
Proga2000}.  with one dominating over the other depending on the shape
of the SED \citep[][strong \civ EQW indicates a
more ionizing SED and large \civ blueshift indicating a less ionizing
SED]{Leighly2004b}.

The Sloan Digital Sky Survey Reverberation Mapping Project
\cite[(SDSS-RM)][]{Shen2015} has a monitored $\sim$350 quasars with
\civ Noting the biases associated with \civ Emission Line Properties
\citep[e.g. increasing systematic offsets with decreasing
signal-to-noise][]{Denney2016}, \citet{Grier2019} report report
significant time delays between the continuum and the CIV 1549
emission line in 52 quasars, and investigate the \civ
radius-luminosity relationship.

%% \citet{Coatman2017} 
%The \civ $\lambda \lambda$1498,1501 broad emission line is visible in optical spectra to redshifts exceeding $z\sim5$. 
\civ has long been known to exhibit significant displacements to the blue and these `blueshifts' almost certainly signal the presence of strong outflows. As a consequence, single-epoch virial black hole (BH) mass estimates derived from \civ velocity widths are known to be systematically biased compared to masses from the hydrogen Balmer lines. 
\citet{Coatman2017} use a large sample of 230 high-luminosity (LBol = 1045.5-1048 erg s-1), redshift $1.5 < z < 4.0$ quasars with both \civ and Balmer line spectra, we have quantified the bias in \civ BH masses as a function of the \civ blueshift. \civ BH masses are shown to be a factor of 5 larger than the corresponding Balmer-line masses at C IV blueshifts of 3000 km s$^{-1}1$ and are overestimated by almost an order of magnitude at the most extreme blueshifts, $\gtrsim$5000 km s$^{-1}$.

@@

\subsection{Line Properties}. 
We use the measured quasar emission line properties from \citet{Shen2011} and \citet{Hamann2017}. \citet{Hamann2017} in particular investigate in robust detail the UV continuum and the \civ (and \nv $\lambda$ $\lambda$1238, 1242) emission lines in over 200,000 quasars in BOSS DR12Q \citep{Paris2017}\footnote{This emission-line catalog can be downloaded from \href{https://datadryad.org/stash/dataset/doi:10.6086/D1H59V}{here}.} The quasar redshift are limited to the range 1.53 ≤ ze ≤ 5.0 so that C IV and the adjacent continuum are covered by BOSS. These measurements provide line profile information and N V/C IV flux ratios...
